
## Building Your First Data Visualization

### Intro to plotting and the ggplot library

We can use the `plot` function to create a scatter plot for the `pressure` data

```{r}
plot(pressure)
```

## Loading Libraries

First, let's load the libraries you will use for this lesson. This is the first thing you should do when writing an R-Markdown document. That way, you ensure that you load all of the necessary libraries prior to running code.

```{r}
# install.packages("ggplot2")
library(ggplot2)
```

Remember to make sure ggplot is loaded into R.

You can do this by running the first chunk of this document, where it has the code `library(ggplot2)`

for ggplot, our first argument will be the dataset `mtcars`

```{r}
ggplot(mtcars)
```

## aes() function for stating your x and y axis

Within the ggplot() function, we'll set up our parameters by using the `aes()` function aes stands for asthetic. For this function, we want to define the x and y axis.

We want to plot how miles per galon `mpg` is related to horsepower `hp`. And then we want to split this up by cylnders `cyl`.

The x-axis will be `hp` and the y-axis will be `mpg`

```{r}
ggplot(mtcars, aes(x = hp, y = mpg))
```

## The power of `+`

## Adding features to your ggplot graph using `+`

We can add new features by using other functions that are part of the ggplot library.

We do this by using the `+` sign

`geom_()` function, which is our geometric object, is used to state the kind of graph we want

Now let's use the `geom_()` family of functions to state what kind of graph we want.

We want a scatterplot, so we are going to use the function `geom_point()`

No arguments are required for `geom_point()`

```{r}
ggplot(mtcars, aes(x = hp, y = mpg)) +
  geom_point()
```

## theme() function for modifying components of your graph

let's set the size of the text by 20 using theme(`text = element_text(size = 20))`

```{r}
ggplot(mtcars, aes(x = hp, y = mpg)) +
  geom_point() +
  theme(text = element_text(size = 20))
```

## labs() function for labeling your graph

```{r}
ggplot(mtcars, aes(x = hp, y = mpg)) +
  geom_point() +
  theme(text = element_text(size = 20)) +
  labs(x = "Horsepower", y = "MPG")
```

## scale_color_gradient() function for using a color gradient on mpg

We want the low mpg to be blue and the high mpg to be red.<br>

IMPORTANT! In order to make this function work, you have to state which variable you want it to color. Let's color `mpg`.

In order to state this, we have to go back to the `aes()` function and write an additional argument. Remember, arguments are separated by `,`

So your aes code should look like this now `aes(x = hp, y = mpg, color = mpg)`

```{r}
ggplot(mtcars, aes(x = hp, y = mpg, color = mpg)) +
  geom_point() +
  theme(text = element_text(size = 20)) +
  labs(x = "Horsepower", y = "MPG") +
  scale_color_gradient(low = "red", high = "green")
```

## facet_wrap() function splitting up the graph by a group

We want to split our graph up by the variable `cyl`<br>

use `~cyl` to wrap by columns

```{r}

```

use `ncol =` to specify how many columns you want in your layout

```{r}

```

