---
title: "Lesson 2"
author: "Fernando Rodriguez"
date: "1/17/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


For this lesson, we'll understand how to develop an organized workflow. 



# 0. Load Libraries
First, we always want to load any relevant libraries we'll be using.
```{r, warning = FALSE}
# install.packages("summarytools")
library(ggplot2) # for plotting
library(summarytools) # for getting data summaries
```
<br><br>

# 1. More workflow basics
Instead of having to use the `insert button` on the top right panel of our 
R-Markdown document, we can use the following code chunks:<br>


<b>alt + command + i (mac)<br>
control + alt + i (windows)</b><br>
<br>

### Dangers of navigating multiple R-Markdown files
One of the main dangers of having multiple R-Markdown files open is that, if you
are using the same variable or dataframe names across both documents, these will get overwritten, depending on what document you are working with.

For instance, the example below, if we have two R-Markdown files that are currently
open, and both have the variable name `a`.

Where, for file 1, a == 4 <br>
and, for file 2, a == 10 <br>

If we run file 2, then a will equal 10 in memory. 
```{r}
# using same variable name
a <- 10
a

```
<br>

### Code comments / commenting out code<br>

Commenting out multiple lines of code<br>
First highlight the code, the run this<br>

control + shift + c (windows and mac)<br>
command + shift + c (mac)<br>
```{r}
# add comments. It doesn't effect the caculations

# a <- 6 + 10
# b <- 10 + 10

```
<br><br>

# 2. Importing Data using `read.csv()`

We are going to use the `read.csv()` function.<br>
There are some pros and cons to using this, but the purpose is to get familiar
with finding full path directories and the consequences of including and excluding
specific arguments. 

Find the path where you stored the `Physics Course Demo Data.csv` file.
Your path should look different from the path below. 

```{r}
demog <- read.csv("/Volumes/GoogleDrive/My Drive/R Workspace/R-for-Ed-Data-Science/Data/Physics Course Demo Data.csv", header = TRUE)
```
<br>

### Getting dataframe structure using `str()` function

```{r}
str(demog)
```
<br><br>

## 3. Vector Types and Classes
R has 6 atomic vector types<br>
Atomic means that it only holds data of a single type<br>
- character: "a" "string values"<br>
- numeric: 2, 1.4, 503<br>
- iteger: 2L (the L tells R to store this as an integer)<br>
- logical: TRUE or FALSE<br>
- complex: 1+4i (complex numbers with real and imaginary parts)<br>
<br><br>

### Factors
Factors are stored as a vector of integer values WITH a cooresponding set of character values.<br>

In many ways factors are a special case of integers.<br>
```{r}
str(demog$agegroup)
```
<br><br>

# 4. Data Reduction (How do you do this?)
Using the `summarytools` library to get overall descriptive info and NAs.<br><br>

However, as you'll see next week, the `dfSummary()` function s is not generating the 
correct Ns. This has to do with how we imported our data. We need to include an
argument that instructs R what to do with our NAs in the .csv file.

```{r}
# view(dfSummary(demog, style = 'rmarkdown'))
```
<br>

Use these additional arguments to have the table display correctly when 
knitting your R-Markdown
```{r}
print(dfSummary(demog, graph.magnif = 0.65), method = 'rmarkdown')
```
<br><br>

### Importing csv files with correct NAs
```{r}
demog2 <- read.csv("/Volumes/GoogleDrive/My Drive/R Workspace/R-for-Ed-Data-Science/Data/Physics Course Demo Data.csv", header = TRUE,
                   na.strings = c ("", NA))
```
<br>

### Re-running dfSummary
```{r}
print(dfSummary(demog2, graph.magnif = 0.65), method = 'rmarkdown')
```