# Project Workflow

For this lesson, we'll understand how to develop an organized workflow. 


## Load Libraries

This is the first thing you want to do when authoring R-Markdown files. For this lesson, we'll use `ggplot2`.

```{r}
library(ggplot2)
```

# More Workflow Basics

## Creating Code Chunks

For mac: alt + command + i

For windows: control + alt + i

```{r}

```

## Creating Comments

use `#` to add comments within code chunks.

```{r}
# I can write anything here about our object a.
a <- 2 + 2 # this is also allowed

```

## Commenting out multiple lines of code

For Mac and Windows: control + shift + c For Mac: command + shift + c

```{r}
a <- 2 + 2
a <- 2 + 2
a <- 2 + 2
a <- 2 + 2
a <- 2 + 2
```



# Importing Data

First, we want to check our working directory getwd()

```{r}
getwd()
```

## using read.csv() to import data

```{r}
# df1 = data frame 1
df1 <- read.csv("/Users/fernanr1/Documents/r-workshop/data/physics-course-data.csv")
```


# Inspecting Data

## Types and Classes

6 atomic vector types (atomic means that an object holds data of a single type) - numeric - integer - string "a" "a word" - logical: TRUE or FALSE - complex 1+4i

## using str() to examine data structure

```{r}
str(df1)
```

## Factors
Factors are variables in R which take on a limited number of different values, variables that are often referred to as categorical variables in the statistics world. 

```{r}
str(df1$agegroup)
```


## use the `View()` function actually look at your data
Note that all functions, like `View()` are case sensitive. Typing `view()` wont' work.
```{r}
View(df1)
```

## Using `summary()` to examine data

```{r}
summary(df1$hsgpa)
```

## Using `table()` to examine raw counts

```{r}
table(df1$hsgpa)
```

# Creating a Histogram
## Using ggplot to creat a histogram of hsgpa
Format is similar to last week's lesson, but here were are using `geom_histogram()` instead of `geom_point()`
```{r}
ggplot(df1, aes(x = hsgpa)) +
  geom_histogram() +
  stat_bin(bins = 50) +
  labs(x = "High School GPA", y = "Number of Students")
```
